The speed of PyPy

### 2010-02-15 talk/pycon2010/pypyspeed/talk.tex
Finish slides

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
\end{frame}

@@

\begin{frame}
  \frametitle{A piece of advice}
  \begin{itemize}
    \item Don't use advanced features if you don't have to
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{xxx}
\end{frame}

>>>>

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{How well it works in practice?}
\end{frame}

@@

\begin{frame}
  \frametitle{A piece of advice}
  \begin{itemize}
    \item don't use advanced features if you don't have to
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Tracing JIT}
  \begin{itemize}
    \item compiler traces the actual execution of Python program
    \item then compiles linear path to assembler
    \item example
    \item mostly for speeding up loops and to certain extent
      recursion
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Removing frame overhead}
  \begin{verbatim}
x = y + z
  \end{verbatim}
  \begin{itemize}
    \item above has 5 frame accesses
    \item they can all be removed (faster!)
      \pause
    \item they prevent optimizations from happening
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Removing object boxing}
  \begin{verbatim}
i = 0
while i < 100:
  i += 1
  \end{verbatim}
  \begin{itemize}
    \item for each iteration we do a comparison and addition
    \item xxx integers on valuestack and xxx integers in locals
    \item all of those can be removed
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Access costs}
  \begin{itemize}
    \item local access costs nothing
    \item global access is cheap, if you don't change global {\ttfamily \_\_dict\_\_} too much XXX rephrase
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Frame escapes}
  \begin{itemize}
    \item calling {\ttfamily sys.\_getframe()}, {\ttfamily sys.exc\_info()}
    \item exception escaping
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Shared dicts (aka hidden classes)}
  \begin{itemize}
    \item instance {\ttfamily \_\_dict\_\_ } lookup becomes a list lookup
    \item if you're evil, it'll bail back to dict lookup
      \pause
    \item only for newstyle classes!
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Version tags}
  \begin{itemize}
    \item dicts on types are version-controlled
    \item this means methods are usually in known places
      \pause
    \item ... if you don't modify them too often
    \item counters on classes are bad
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Call costs}
  \begin{itemize}
    \item calls can be inlined
    \item simple arguments are by far the best
    \item avoid {\ttfamily *args} and {\ttfamily **kwds}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Allocation patterns}
  \begin{itemize}
    \item PyPy uses a moving GC (like JVM, .NET, etc.)
    \item pretty efficient for usecases with a lot of
      short-living objects
    \item objects are smaller than on CPython
    \item certain behaviors are different than on CPython
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Differencies}
  \begin{itemize}
    \item no refcounting semantics
    \item {\ttfamily id(obj)} can be expensive as it's a complex
      operation on a moving GC
    \item a large list of new objects is a bad case behavior
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{General rules}
  \begin{itemize}
    \item don't try to outsmart your compiler
    \item simple is better than complex
    \item metaprogramming is your friend
    \item measurment is the only meaningful way to check
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Problems}
  \begin{itemize}
    \item long traces - tracing is slow
    \item megamorphic calls
    \item metaclasses
    \item class global state
      \pause
    \item years of optimizations against CPython
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Future}
  \begin{itemize}
    \item release end March
    \item to try it out
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{That's all!}
  \begin{itemize}
    \item Q \& A
    \item http://morepypy.blogspot.com
    \item http://pypy.org
    \item http://merlinux.eu
  \end{itemize}
\end{frame}
